### 2021-07-07 rad-icfp-2021/correctness.tex
fixing boundary issues

    \begin{tabular}{r c l}
    $\seml \reals \semr$ & $\defeq$& $\RR$ \\
    $\seml$!T1xT2!$\semr$ & $\defeq$& $\seml$!T1!$\semr \times\seml$!T2!$\semr$ \\
    $\seml \Array{\reals}{n}\semr$ & $\defeq$ & $\prod_{1\leq i \leq n} \RR$ 
    \end{tabular}
    \begin{tabular}{r c l}
        !$\seml$op1$\semr$! & $\defeq$ & op1: $\reals\to\reals$ \\
        !$\seml$op2$\semr$! & $\defeq$ & op2: $\reals\times \reals\to\reals$ \\
        !$\seml$c$\semr$! & $\defeq$ & $c\in\RR$ \\
        !$\seml$map2 (x,y.e$_1$) e$_2$ e$_3\semr_\gamma$ ! & $\defeq$ & \\
        !$\seml$reduce (x,y.e$_1$) e$_2$ e$_3\semr_\gamma$! & $\defeq$ &
    \end{tabular}

@@

\begin{definition}[Mutlicat from concat]
    TODO
\end{definition}

@@

\[
\begin{tikzcd}
    SourceUNF \ar[r,"\sem{-}"] & Concat1 \ar[r,"\Dsynrevsymbol_{\rho}"] & Concat2 \ar[r,"\UNFSymbol^{-1}"] & C_{Target} \\
    Source \ar[r,"\UNFSymbol"] & C_{Source} \ar[u, hook] & TargetUNF \ar[u,"\sem{-}"] & Target \ar[u, hook] 
\end{tikzcd}
\]

>>>>

    \begin{tabular}{r c l}
    $\seml \reals \semr$ & $\defeq$& $\RR$ \\
    $\seml$!T1xT2!$\semr$ & $\defeq$& $\seml$!T1!$\semr \times\seml$!T2!$\semr$ \\
    $\seml \Array{\reals}{n}\semr$ & $\defeq$ & $\prod_{1\leq i \leq n} \RR$ \\
    !$\seml$op1$\semr$! & $\defeq$ & op1: $\reals\to\reals$ 
    \end{tabular}
    \begin{tabular}{r c l} 
        !$\seml$op2$\semr$! & $\defeq$ & op2: $\reals\times \reals\to\reals$ \\
        !$\seml$c$\semr$! & $\defeq$ & $c\in\RR$ \\
        !$\seml$map2 (x,y.e$_1$) e$_2$ e$_3\semr_\gamma$ ! & $\defeq$ & TODO \\
        !$\seml$reduce (x,y.e$_1$) e$_2$ e$_3\semr_\gamma$! & $\defeq$ &
    \end{tabular}

@@

\begin{definition}[Mutlicat from concat]
    We define $\CSource$ to be the multicategory with the same objects as Source and 
    $\CSource([A_1,\ldots,A_n],B)=\ConcatS([A_1,\ldots,A_n],[A_1,\ldots,A_n,B])$.
\end{definition}

@@

\[
\begin{tikzcd}
    SourceUNF \ar[r,"\sem{-}"] & Concat1 \ar[r,"\Dsynrevsymbol_{\rho}"] & Concat2 \ar[r,"\UNFSymbol^{-1}"] & C_{Target} \\
    Source \ar[r,"\UNFSymbol"] & \CSource \ar[u, hook] & TargetUNF \ar[u,"\sem{-}"] & Target \ar[u, hook] 
\end{tikzcd}
\]

### 2021-07-07 rad-icfp-2021/figures/macro_diff.tex
fixing boundary issues

\begin{figure*}[t]
    \begin{tabular}{r c l}
    $\Dsynrevsymbol$(!var!$_{\Gamma,i}$) &=& !<var!$_{\Gamma,i}$!, J!$^T$!var!$_{\Gamma,i}$!>! \\
    $\Dsynrevsymbol$(!op!$_{\Gamma,n}$) &=& !<op!$_{\Gamma,n}$!, J!$^T$!op!$_{\Gamma,n}$!>! \\ 
    $\Dsynrevsymbol$(!e1;e2!) &=& $\Dsynrevsymbol$(!e1!);$\Dsynrevsymbol$(!e2!)\\ 
    $\Dsynrevsymbol$(!map!$_{x.e}$) &=& !<map!$_{x.e}$!, J!$^T$!map!$_{x.e}$!>! \\ 
    $\Dsynrevsymbol$(!map2!$_{x,y.e}$) &=& !<map2!$_{x,y.e}$!, J!$^T$!map2!$_{x,y.e}$!>! \\ 
    $\Dsynrevsymbol$(!foldl!$_{x,y.e}$) &=& !<foldl!$_{x,y.e}$!, J!$^T$!foldl!$_{x,y.e}$!>! \\  
    \end{tabular}
    \caption{Reverse-mode differentiation from Source UNF to Target UNF}
    \label{fig:diff_macro}    \end{figure*}

>>>>

\begin{figure*}[t]
    \begin{tabular}{r c l}
        $\Dsynrevsymbol_{\rho}$(!A1,$\ldots$,An!) &=& !A1,$\ldots$,An,(A1x$\ldots$xAn)->!$\rho$
    \end{tabular}

    \medskip
    \begin{tabular}{r c l}
    $\Dsynrevsymbol_{\rho}$(!var!$_{T;i}$) &=& $F$(!var$_{T;i}$, J$^T$var$_{T;i}$!) \\
    $\Dsynrevsymbol_{\rho}$(!op!$_{\Gamma;n}$) &=& $F$(!op$_{T;n}$,J$^T$op$_{T;,n}$!) \\ 
    $\Dsynrevsymbol_{\rho}$(!pair!$_{T;A\times B}$) &=& $F$(!pair$_{T;A\times B}$, J$^T$pair$_{T;A\times B}$!) \\
    $\Dsynrevsymbol_{\rho}$(!proj!$_{T1;T2;T3}$) &=& $F$(!proj$_{T1;T2;T3}$,J$^T$proj$_{T1;T2;T3}$!) \\
    $\Dsynrevsymbol_{\rho}$(!e1!$\comp$!e2!) &=& $\Dsynrevsymbol_{\rho}$(!e1!); $\Dsynrevsymbol_{\rho}$(!e2!)\\ 
    $\Dsynrevsymbol_{\rho}$(!map2!$_{T;x,y.e}$) &=& $F$(!map2$_{T;x,y.e}$, J$^T$map2$_{T;x,y.e}$!) \\
    $\Dsynrevsymbol_{\rho}$(!reduce!$_{T;x,y.e1;e2}$) &=& $F$(!reduce$_{T;x,y.e1;e2}$, J$^T$reduce$_{T;x,y.e1;e2}$!) \\
    \end{tabular}\\
    where !$F$(A,B)$\defeq$ <proj$_{T;\underline{T}->\rho;[]} \comp$ A, proj$_{[];T;\underline{T}->\rho} \icomp$ (proj$_{T;\underline{T}->\rho;[]}$ $\comp$ B)>!
    \caption{Reverse-mode differentiation from Source UNF to Target UNF}
    \label{fig:diff_macro}    
\end{figure*}

### 2021-07-07 rad-icfp-2021/generalization.tex
minor
fixing boundary issues

\section{Generalizations}
\label{sec:generalization}

@@

We can now compute the reverse-mode transformation as follows.
\begin{center}
\begin{tabular}{r c l}
$\directD{\rho}{\Gamma}{Y}$($\Gamma\vdash $ !reduce (x,y.e1) e2 e3!) 
&=& $\ldots$ as before up until !A3! \\
&& !let B0 = map2 (a,b ->(!$\grad_\Gamma$!e)[a/x,b/y] A0 A) in!\\
&& !let B1 = scanr !$1_\Gamma$ $\widehat{\times}$ !B0 in! \\
&& !let B2 = reduce !$\widehat{+}$ $0_\Gamma$! B1 in! \\
&& !<reduce (x,y.e1) y1 A, fun (x1,$\ldots$,xn,z) ->! \\
&& !Y((x$_1$,$\ldots$,x$_n$)!$\widehat{+}$!B2,0,map2 (a,b. a*b*z) A2 A3)!
\end{tabular}
\end{center}

\subsection{Adding more array operators}
\label{sub:Adding more array operators}

@@

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$($\Gamma\vdash $ !foldl (x,y.e$_1$) e$_2$ e$_3$!) &=&
            !let v,Y$_1$ = $\directD{\rho}{\Gamma}{Y}$(e$_2$) in! \\
            && !let A,Y$_2$ = $\directD{\rho}{\Gamma,v}{Y_1}$(e$_3$) in! \\
            && !let A$_0$,r$_1$ = (scanl (x,y.e$_1$) v A) in! \\
            && !let A$_1$ = map2 (a,b.(!$\grad_{\{x\}}$!e$_1$)[a/x,b/y]) A$_0$ A! \\
            && !let A$_2$ = map2 (a,b.(!$\grad_{\{y\}}$!)e$_1$[a/x,b/y]) A$_0$ A! \\
            && !let r$_2$, A$_3$ = scanr * 1 A$_1$! \\
            && !<r$_1$, fun (x$_1$,$\ldots$,x$_n$,z) ->! \\
            && !let y$_1$,B = r$_2$*z ,map2 (x,y. x*y*z) A$_2$ A$_3$ in! \\
            && !Y$_2$(x$_1$,$\ldots$,x$_n$,y$_1$,B)>! \\
\end{tabular}
\end{center}

@@

\subsection{Adding more scalar operators} % (fold)
\label{sub:Adding more scalar operators}

@@

\subsection{Adding conditionals} % (fold)
\label{sub:Adding conditionals}

@@

\begin{tabular}{r c l}
$\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) &=& !if e1 then !$\directD{\rho}{\Gamma}{Y}$!(e2) else !$\directD{\rho}{\Gamma}{Y}$!(e3)! 
\end{tabular}

@@

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) 
    &=& !let b=e1 in!   \\
    && !<if b then e2 else e3, fun (x1,$\ldots$,xn,z) ->! \\
    && !Y(b*!$\directD{\rho}{\Gamma}{Y}$!(e2)(x1,$\ldots$,xn,z)+!\\
    && \quad!(1-b)*!$\directD{\rho}{\Gamma}{Y}$!(e3)(x1,$\ldots$,xn,z))>!
\end{tabular}

@@

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) 
    &=& !let b=e1 in!   \\
    && !<if b then e2 else e3, fun (x1,$\ldots$,xn,z) ->! \\
    && !Y((x1,$\ldots$,xn)!$\widehat{+}$($\grad_\Gamma$!(e2)*b!$\widehat{+}\grad_\Gamma$!(e3)*(1-b))*z)>!\\
\end{tabular}

@@

\subsection{Adding general arrays} % (fold)
\label{sub:Adding general arrays}

@@

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$($\Gamma\vdash $ !map2 (x,y.e1: G) e2 e3!) &=&  
    !let A,Y1 = !$\directD{\rho}{\Gamma}{Y}$!(e2) in! \\
    && !let B,Y2 = !$\directD{\rho}{\Gamma,A}{Y1}$!(e3) in! \\
    && !let C=!$\grad_{\Gamma}$!e1 in!\\
    && !<map2 (x,y.e1) A B, fun (x1,$\ldots$,xn,Z) -> !\\
    && !Y2( (x1,$\ldots$,xn)!$\widehat{+}$!(C!$\widehat{*}$!(reduce + 0 Z)),!\\
    && \quad\quad! map2 (a,b.(!$\grad_{\{x\}}$!e1)[a/x]!$\widehat{*}$!b) A Z,!\\
    && \quad\quad! map2 (a,b.(!$\grad_{\{y\}}$!e1)[a/x]!$\widehat{*}$!b) B Z )>!\\
    \end{tabular}
    \end{center}

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!reduce (x,y.e$_{1}$) e$_{2}$ e$_{3}$!) &=&
    !let y$_{1}$,Y$_{1}$ = !$\directD{\rho}{\Gamma}{Y}$!(e$_{2}$) in! \\
    && !let A,Y$_{2}$ = !$\directD{\rho}{\Gamma,y_1}{Y_1}$!(e$_{3}$) in! \\
    && !let A$_{0}$ = shift1R (scanl (x,y.e$_{1}$) y$_{1}$ A) in! \\
    && !let A$_{1}$ = shift1L (map2 (a,b.(!$\grad_{\{x\}}$!e$_{1}$)[a/x,b/y]) A$_{0}$ A) in! \\
    && !let A$_{2}$ = map2 (a,b.(!$\grad_{\{y\}}$!)e$_{1}$[a/x,b/y]) A$_{0}$ A in! \\
    && !let A$_{3}$ = scanr $\widetilde{*}$ 1 A$_{1}$ in! \\
    && !<reduce (x,y.e$_{1}$) y$_{1}$ A, fun (x$_{1}$,$\ldots$,x$_n$,z) ->! \\
    && !Y$_{2}$(x$_{1}$,$\ldots$,x$_n$, map2 (x,y. x$\widehat{*}$(y$\widehat{*}$z)) A$_{2}$ A$_{3}$>! \\
\end{tabular}
\end{center}

>>>>

\section{Beyond the Source language}
\label{sec:generalization}

@@

We can now compute the reverse-mode transformation as follows.
\begin{center}
\begin{tabular}{r c l}
$\directD{\rho}{\Gamma}{Y}$(!reduce (x,y.e1) e$_2$ e$_3$!) 
&=& $\ldots$ as before up until !A_$3$! \\
&& !let B$_0$ = map2 (a,b ->(!$\grad_\Gamma$!e)[a/x,b/y] A$_0$ A) in!\\
&& !in let B$_1$ = scanr $1_\Gamma$ $\widehat{\times}$ B$_0$ in! \\
&& !let B$_2$ = reduce $\widehat{+}$ $0_\Gamma$ B$_1$ in! \\
&& !<reduce (x,y.e1) y$_1$ A, fun (x$_1$,$\ldots$,x$_n$,z) ->! \\
&& !Y((x$_1$,$\ldots$,x$_n$)$\widehat{+}$B$_2$,0,map2 (a,b. a*b*z) A$_2$ A$_3$)!
\end{tabular}
\end{center}

\subsection{More array operators}
\label{sub:Adding more array operators}

@@

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!foldl (x,y.e$_1$) e$_2$ e$_3$!) &=&
            !let v,Y$_1$ = $\directD{\rho}{\Gamma}{Y}$(e$_2$) in! \\
            && !let A,Y$_2$ = $\directD{\rho}{\Gamma,v}{Y_1}$(e$_3$) in! \\
            && !let A$_0$,r$_1$ = (scanl (x,y.e$_1$) v A) in! \\
            && !let A$_1$ = map2 (a,b.(!$\grad_{\{x\}}$!e$_1$)[a/x,b/y]) A$_0$ A! \\
            && !let A$_2$ = map2 (a,b.(!$\grad_{\{y\}}$!)e$_1$[a/x,b/y]) A$_0$ A! \\
            && !let r$_2$, A$_3$ = scanr * 1 A$_1$! \\
            && !<r$_1$, fun (x$_1$,$\ldots$,x$_n$,z) ->! \\
            && !let y$_1$,B = r$_2$*z ,map2 (x,y. x*y*z) A$_2$ A$_3$ in! \\
            && !Y$_2$(x$_1$,$\ldots$,x$_n$,y$_1$,B)>! \\
\end{tabular}
\end{center}

@@

\subsection{Non-smooth scalar operators} % (fold)
\label{sub:Adding more scalar operators}

@@

\subsection{Conditionals} % (fold)
\label{sub:Adding conditionals}

@@

\begin{tabular}{r c l}
$\directD{\rho}{\Gamma}{Y}$(!if e$_1$ then e$_2$ else e$_3$!) &=& !if e$_1$ then $\directD{\rho}{\Gamma}{Y}$(e$_2$) else $\directD{\rho}{\Gamma}{Y}$(e$_3$)! 
\end{tabular}

@@

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!if e$_1$ then e$_2$ else e$_3$!) 
    &=& !let b=e$_1$ in!   \\
    && !<if b then e$_2$ else e$_3$, fun (x$_1$,$\ldots$,x$_n$,z) ->! \\
    && !Y(b*!$\directD{\rho}{\Gamma}{Y}$!(e$_2$)(x$_1$,$\ldots$,x$_n$,z)+!\\
    && \quad!(1-b)*!$\directD{\rho}{\Gamma}{Y}$!(e$_3$)(x$_1$,$\ldots$,x$_n$,z))>!
\end{tabular}

@@

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!if e$_1$ then e$_2$ else e$_3$!) 
    &=& !let b=e$_1$ in!   \\
    && !<if b then e$_2$ else e$_3$, fun (x$_1$,$\ldots$,x$_n$,z) ->! \\
    && !Y((x$_1$,$\ldots$,x$_n$)!$\widehat{+}$($\grad_\Gamma$!(e$_2$)*b!$\widehat{+}\grad_\Gamma$!(e$_3$)*(1-b))*z)>!\\
\end{tabular}

@@

\subsection{General arrays} % (fold)
\label{sub:Adding general arrays}

@@

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!map2 (x,y.e$_1$: G) e$_2$ e$_3$!) &=&  
    !let A,Y$_1$ = !$\directD{\rho}{\Gamma}{Y}$!(e2) in! \\
    && !let B,Y$_2$ = !$\directD{\rho}{\Gamma,A}{Y1}$!(e$_3$) in! \\
    && !let C=!$\grad_{\Gamma}$!e$_1$ in!\\
    && !<map2 (x,y.e$_1$) A B, fun (x$_1$,$\ldots$,x$_n$,Z) -> !\\
    && !Y$_2$( (x$_1$,$\ldots$,x$_n$)!$\widehat{+}$!(C!$\widehat{*}$!(reduce + 0 Z)),!\\
    && \quad\quad! map2 (a,b.(!$\grad_{\{x\}}$!e$_1$)[a/x]!$\widehat{*}$!b) A Z,!\\
    && \quad\quad! map2 (a,b.(!$\grad_{\{y\}}$!e$_1$)[a/x]!$\widehat{*}$!b) B Z )>!\\
    \end{tabular}
    \end{center}

\begin{center}
\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}$(!reduce (x,y.e$_{1}$) e$_{2}$ e$_{3}$!) &=&
    !let y$_{1}$,Y$_{1}$ = !$\directD{\rho}{\Gamma}{Y}$!(e$_{2}$) in! \\
    && !let A,Y$_{2}$ = !$\directD{\rho}{\Gamma,y_1}{Y_1}$!(e$_{3}$) in! \\
    && !let A$_{0}$ = shift1R (scanl (x,y.e$_{1}$) y$_{1}$ A) in! \\
    && !let A$_{1}$ = shift1L (map2! \\ 
    && !      (a,b.(!$\grad_{\{x\}}$!e$_{1}$)[a/x,b/y]) A$_{0}$ A) in! \\
    && !let A$_{2}$ = map2! \\
    && !      (a,b.(!$\grad_{\{y\}}$!)e$_{1}$[a/x,b/y]) A$_{0}$ A in! \\
    && !let A$_{3}$ = scanr $\widetilde{*}$ 1 A$_{1}$ in! \\
    && !<reduce (x,y.e$_{1}$) y$_{1}$ A, fun (x$_{1}$,$\ldots$,x$_n$,z) ->! \\
    && !Y$_{2}$(x$_{1}$,$\ldots$,x$_n$, map2 (x,y. x$\widehat{*}$(y$\widehat{*}$z)) A$_{2}$ A$_{3}$>! \\
\end{tabular}
\end{center}

### 2021-07-07 rad-icfp-2021/purerad.tex
fixing boundary issues

\begin{tabular}{l}
    !fun (x$_{1}$,$\ldots$,x$_n$) -> let y$_{1}$,$\ldots$,y$_n$ = e$_1$,$\ldots$,e$_n$ in (f$_{1}$,$\ldots$,f$_n$)!  $\transto$ \\
    !fun (x$_{1}$,$\ldots$,x$_n$) -> let y$_{n-1}$,y$_n$ =e$_{n-1},$e$_1$ in (f$_{1}$[x$_{1}$/y$_{1}$],$\ldots$,f$_{n-2}[$x$_{n-2}$/y$_{n-2}]$,f$_{n-1}$,f$_{n}$)!
\end{tabular}
It means that it does not change the evaluation cost of most of the !f!$_{i}$ for $1\leq i \leq n-2$.
The new evaluation cost is reduced to the sum of the cost of evaluating 
the !f!$_{i}$ in addition to the cost of evaluating !e!$_{n-1}$ and !e!$_{n}$, O($n$) movement of variables.

>>>>

\begin{tabular}{l}
    !fun (x$_{1}$,$\ldots$,x$_n$) -> let y$_{1}$,$\ldots$,y$_n$ = e$_1$,$\ldots$,e$_n$ in (f$_{1}$,$\ldots$,f$_n$)!  $\transto$ \\
    !fun (x$_{1}$,$\ldots$,x$_n$) -> let y$_{n-1}$,y$_n$ = e$_{n-1},$e$_1$ in!\\
    \hspace{3cm}!(f$_{1}$[x$_{1}$/y$_{1}$],$\ldots$,f$_{n-2}[$x$_{n-2}$/y$_{n-2}]$,f$_{n-1}$,f$_{n}$)!
\end{tabular}

It means that it does not change the evaluation cost of most of the !f!$_{i}$ for $1\leq i \leq n-2$.
The new evaluation cost is reduced to the sum of the cost of evaluating 
the !f!$_{i}$ in addition to the cost of evaluating !e!$_{n-1}$ and !e!$_{n}$, gaining O($n$) movement of variables.

### 2021-07-07 rad-icfp-2021/unf.tex
fixing boundary issues

\begin{example}
    !cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$! 
    represents !x$_1$:$\reals$, x$_2$:$\reals$, x$_3$:$\reals$ $\vdash$ let (x$_1$,x$_2$,x$_3$,x$_4$)=(x$_1$,x$_2$,x$_3$,cos(x$_3$)) in (x$_1$,x$_2$,<x$_3$,x$_4$>)!.
    $\comp$ can more generally be understood as a let binding of a tuple, and primitives only act on the last parts of the tuple, always returning the context.
\end{example}

@@

\subsection{transformations to and from UNF} % (fold)
\label{sub:transformations to and from UNF}

@@

\begin{lemma}[Well typedness $\UNFSymbol$]
    If !$\Gamma$=x$_1$:A$_{1}$,$\ldots$,x$_n$:A$_{n}$ $\vdash$ e:B! then
   !A$_{1}$,$\ldots$,A$_{n}$ $\vdash$ $\UNFSymbol$(e): A$_{1}$,$\ldots$,A$_{n}$,B!.
\end{lemma}

@@

Target UNF does not have variables. 
A context type !T=[A$_1$,$\ldots$,A$_n$]! is transformed to a context !$\Gamma$=x$_1$:A$_1$,$\ldots$,x$_n$:A$_n$!.
We use this convention when definiting $\UNFSymbol^{-1}$. 
For operators like !op$_{\Gamma;m}$!, we extend the notation above by saying that the canonical variables are 
!x$_1$:A$_1$,$\ldots$,x$_n$:A$_n$,x$_{n+1}$:$\reals$,$\ldots$,x$_{n+m}$:$\reals$!.

>>>>

\begin{example}
    The term in Source UNF !cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$! 
    intuitively represents a term \\
    !x$_1$:$\reals$, x$_2$:$\reals$, x$_3$:$\reals$ $\vdash$ let (x$_1$,x$_2$,x$_3$,x$_4$)=(x$_1$,x$_2$,x$_3$,cos(x$_3$)) in (x$_1$,x$_2$,<x$_3$,x$_4$>)!.
    The operator $\comp$ can more generally be understood as a let binding of a tuple, and primitives only act on the last parts of the tuple, always also returning their whole context.
\end{example}

@@

\subsection{Transformations to and from UNF} % (fold)
\label{sub:transformations to and from UNF}

@@

\begin{lemma}[Well typedness $\UNFSymbol$]
    Let !$\Gamma$=x$_1$:A$_{1}$,$\ldots$,x$_n$:A$_{n}$ $\vdash$ e:B! be a term in Source. Then
    \begin{center}
        !A$_{1}$,$\ldots$,A$_{n}$ $\vdash$ $\UNFSymbol$(e): A$_{1}$,$\ldots$,A$_{n}$,B!.
    \end{center}
\end{lemma}

@@

Target UNF does not have variables. 
A context type !T=[A$_1$,$\ldots$,A$_n$]! is transformed to a context \\
!$\Gamma$=x$_1$:A$_1$,$\ldots$,x$_n$:A$_n$!.
We use this convention when definiting $\UNFSymbol^{-1}$. 
For operators like !op$_{\Gamma;m}$!, we extend the notation above by saying that the variables from the context are \\
!x$_1$:A$_1$,$\ldots$,x$_n$:A$_n$,x$_{n+1}$:$\reals$,$\ldots$,x$_{n+m}$:$\reals$!.
